CSD, Disjoint components and trayectory detection

### 2019-10-15 ArticuloCSDyCM/Paper03.tex
talacha auxiliar
bla bla

\documentclass[12pt]{article}

@@

\doublespacing
\linenumbers

@@

Evoked activity by electrical stimulus yields simpler and cleaner patterns in the LFP
measurements. This facilitates the exemplification of our analysis, so we shall
start the exposition of the usage with these kind of data.
The data was obtained by stimulating certain specific points on the dentate gyrus
of the slice of hippocampus with different voltage intensities. Details of the experiment
have been published elsewhere \cite{Franco2018}. We shall borrow some of
the experimental data for illustrating our purposes.

@@

The LFP is averaged over three events, produced by three successive
electrical stimuli in the same point at constant voltage. The
response of the cells is almost identical in each event, which
allows us to use averages as a denoising mechanism. After averaging
the synchronized data we perform kCSDA, as in this case we end up
with many saturated electrodes and dCDSA is not viable.
In the figure \ref{lfpevocada} we show some frames of the post-stimulus
activity as a heat map, along
with a diagram of the somatic layers and the
stimulus point. The saturated area is quite large, and covers
a large section of the GD structure. In this case we shall
restrict the CM track analysis to the respondent area in CA3.

@@

The figure \ref{lfpcsdcm}  compares the representation of the activity in
the LFP, CSD and CM spaces. The CSD
representation provides better spatial localization of active sites, and the
CM serves as better dynamical tracker. The figure illustrates
that point by choosing the time interval of response in CA3 after the stimulus.
CM representation permits us to track the activity with less
interference from the saturated region. The small, clearly localized centers of
mass separate clearly from the big patch over the saturated region.
In the figure \ref{lfpcsdcmsub} we repeat the procedure but focusing in
the r.o.i., which includes CA3 str. pyr. We take advantage of the
detailed frame to show the borders of the disconnected sets that
we have chosen for the CM analysis in the right column.

In the figure \ref{evotracks} we concatenate the successive centers of mass
for both sinks and sources and show how the anterograde reaction is
very well localized over the Pyramidal Stratum at precise moments after
stimulus. This dataset provides us with a very clean example of the
trajectory analysis. The very well localized activity and coherent
action gives us trajectories that are in accordance with the canonical
source and sink distribution, but adding the temporal dynamical displacement
to it. We have been very thorough with this example as it will help us
to understand the more complicated data that shall be presented in
the next examples.

@@

\subsection{Application to epileptic-like activity in the hippocampus}
We performed first the analysis to the most complicated data,
numerically and qualitatively, which comes from a series of epileptic burst
that occurred after 15 minutes of application of 4AP to living slices of
hippocampus. Each burst occurs with activity spreading over many sites almost
simultaneously. This turned out to be a test of the applicability and
robustness of our method. The necessity of having a good error value for the inactive set was more apparent here than anywhere else. Also the most striking features of the propagation of activity, coded now as trajectories of the CM, could be seen here, as the epileptic burst covered the whole CA structure. A sample of the raw data in $\mu V$ can be seen in the figure \ref{lfpycsd4ap}. Five key moments in the epileptic burst have been selected, one just before the onset of the attack, the next one at the peak and one  following it, and the last at the subsiding moments. The figure compares the Raw data
in $\muV{}$ in the first row, then the crude CSD representation in the second row, and
in the third raw a CSD after a Gaussian Blur Filter has been applied,
with a $\sigma$ value of $ \mum{42}$. After the Gaussian Blur acts a sort of
spatial denoising mechanism, the disjoint components are apparent to the eye.
In the figure \ref{csdfrontera} we apply the border criterion for the disconnected
components. Again, in the first row we have the CSD without the Gaussian Blur.
Even though well localized patches of activity are to be seen over the
stratum pyramidale, large parts of the activity remain hidden by the noise.
After the Gaussian Blur has been applied the joint sets of activity are evident.

In the figure \ref{tray4ap} we show the attack on the trajectories representation. Fine lines
connect the blue (sink) and red (source) dots, that indicate the trajectories along various activation
sites around the whole CA structure, and part of GD. Very small trajectories precede
the onset of the epileptiform attack in CA1. At around 60ms after the 4AP dose a burst
of localized trajectories appear near CA2. By the 70ms mark, the activity has spread
retrograde through most of the CA structure. Between the 80 and 90 ms marks
GD shows knot-like trajectories. By 140ms the remaining activity is in the proximal
area of CA1.

@@

bla bla.

>>>>

\documentclass[12pt, letterpaper]{article}

@@

% \doublespacing
% \linenumbers

@@

Evoked activity by electrical stimulus yields simpler and cleaner patterns in the LFP
measurements. This facilitates the exemplification of our analysis, and gives us
a good starting case. It comes at a cost, nonetheless, as many of the electrodes
remain saturated after the stimulus and well beyond the measurement interval.
The data was obtained by stimulating  specific points on the dentate gyrus
in the slice of hippocampus with different voltage intensities. Details of the experiment
have been published elsewhere \cite{Franco2018}. We shall borrow some of
the experimental data for our purposes.

@@

The LFP is averaged over three events, produced by three successive
electrical stimuli in the same point at constant voltage. The
response of the cells is almost identical in each event, which
allows us to use averages as a denoising mechanism. After averaging
we perform kCSDA. In this case there are
many saturated electrodes and dCDSA is not viable.
In the figure \ref{lfpevocada} we show some frames of the post-stimulus
activity as a color map, along
with a diagram of the somatic layers and the
stimulus point. The saturated area is quite large, and covers
a large section of the GD structure. In this case we shall
restrict the CM track analysis to the respondent area in CA3.

@@

The figure \ref{lfpcsdcm}  compares the representation of the activity in
the LFP, CSD and CM spaces. The CSD
representation provides better spatial localization of active sites, but the
CM serves as better dynamical tracker. The figure illustrates
that point by choosing the time interval of response in CA3 after the stimulus.
CM representation permits us to track the activity with less
interference from the saturated region. The small, clearly localized centers of
mass separate clearly from the big patch over the saturated region.
In the figure \ref{lfpcsdcmsub} we repeat the procedure but focusing in
the r.o.i., which includes CA3 str. pyr. We take advantage of the
detailed frame to show the borders of the disconnected sets that
we have chosen for the CM analysis in the right column.

In the figure \ref{evotracks} we concatenate the successive centers of mass
for both sinks and sources and show how the anterograde reaction is
very well localized over the Pyramidal Stratum at a precise moment after
stimulus. This dataset provides us with a very clean example of the
trajectory analysis. The localized activity and coherent
action gives us trajectories that are in accordance with the expected
source and sink distribution, but adding the temporal dynamical displacement
to it. We have been very thorough with this example as it will help us
to understand the more complicated data that shall be presented in
the next examples.

@@

\subsection{Application to epileptic-like activity in the hippocampus}
This is probably the most complicated data, 
numerically and qualitatively, which comes from a series of epileptic burst
that occurred after 15 minutes of application of 4AP to living slices of
hippocampus. Each burst occurs with activity spreading over many sites almost
simultaneously. This turned out to be a test of the applicability and
robustness of our method.
The necessity of having a good error value for the inactive set was more apparent here
than in the other cases. Also, the most striking features of the propagation of activity,
coded now as trajectories of the CM, could be seen here, as the epileptic burst covered the whole
CA structure. A sample of the raw data in $\muV{}$ can be seen in the figure \ref{lfpycsd4ap}.
Five key moments in the epileptic burst have been selected, one just before the onset of the attack, the next one at the peak and one  following it, and the last at the subsiding moments. The figure compares the Raw data
in $\muV{}$ in the first row, then the raw CSD representation in the second row, and
in the third raw a CSD after a Gaussian Blur Filter has been applied,
with a $\sigma$ value of $ \mum{42}$. After the Gaussian Blur acts a sort of
spatial denoising mechanism, the disjoint components are apparent to the eye.
In the figure \ref{csdfrontera} we apply the border criterion for the disconnected
components. Again, in the first row we have the CSD without the Gaussian Blur.
Even though well localized patches of activity are to be seen over the
stratum pyramidale, large parts of the activity remain hidden by the noise.
After the Gaussian Blur has been applied the joint sets of activity are evident.

In the figure \ref{tray4ap} we show the epileptic burst on the trajectories representation.
Fine lines
connect the blue (sink) and red (source) dots, that indicate the trajectories
along various activation
sites around the whole CA structure, and part of GD. Very small trajectories precede
the onset of the epileptiform attack in CA1. At around 60ms after the 4AP dose a burst
of localized trajectories appear near CA2. By the 70ms mark, the activity has spread
retrograde through most of the CA structure. Between the 80 and 90 ms marks
GD shows knot-like trajectories. By 140ms the remaining activity is in the proximal
area of CA1.

@@
